---
title: "Milestone 4"
author: "Meredith Klashman and Ruben Vargas"
date: "11/29/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
```

# Data Connection and Cleanup from Milestone 3 
```{r}
getwd()

countydemo <- read.csv("/home/rstudio/PHW251_MK_RV/ca_county_demographics.csv",
                      header=TRUE %>%
  select(-c("X", "families", "hse_units", "ave_fam_sz", "vacant", "county_fips")) %>%
  rename(County = name))

mortality <- read.csv("/home/rstudio/PHW251_MK_RV/mort_by_county.csv", header = 
                       TRUE)
hospital_cost <- read.csv("/home/rstudio/PHW251_MK_RV/oshpd_hospital_cost.csv", 
                        header = TRUE)

# countydemo <- read.csv("/Users/rubenvargas/Desktop/Public Health W251/PHW251_MK_RV/ca_county_demographics.csv",
                       # header = T) %>%
  # select(-c("X", "families", "hse_units", "ave_fam_sz", "vacant", "county_fips")) %>%
  # rename(County = name)

# mortality <- read.csv("/Users/rubenvargas/Desktop/Public Health W251/PHW251_MK_RV/mort_by_county.csv",
                       # header = T)

# hospital_cost <- read.csv("/Users/rubenvargas/Desktop/Public Health W251/PHW251_MK_RV/oshpd_hospital_cost.csv",
                      # header = T)
```

```{r}
#Edit Mortality to replace NA with 0 

mortality[is.na(mortality)] = 0 
```

```{r}
#Create a Variable to Indicate Chronic Condition 
 
#Chronic Health Conditions defined as the following: 
    #Alzheimer's disease
    #Chronic lower respiratory diseases
    #Diabetes mellitus
    #Essential hypertension and hypertensive renal disease
    #Chronic liver disease and cirrhosis
    #Parkinson's disease

mortality <- mortality %>% 
  mutate(Chronic = case_when(Cause_Desc %in% c("Alzheimer's disease",
                    "Chronic lower respiratory diseases", "Diabetes mellitus",
                    "Essential hypertension and hypertensive renal disease",
                    "Chronic liver disease and cirrhosis", 
                    "Parkinson's disease") ~ "Yes",
                    TRUE ~ "No"))
```

```{r}
#subtotal for mortality from chronic in each county per year 

mortality_chronic <- mortality %>% 
              filter(Chronic == "Yes")%>%
              filter(Strata == "Total Population") %>%
              group_by(County,Year) %>%
              mutate(County_Year_Chronic_Mortality = sum(Count))

```

```{r}
#Calculate Rent vs Homeowners Ratio 

countydemo <- countydemo %>% 
              mutate(RatioRenttoOwn = renter_occ/owner_occ)

```

```{r}
#locate 5 counties that share three common attributes: low population per 
#square mile `pop12_sqmi1`, high median age 
#`med_age`, a high proportion of renters vs. homeowners 

countydemo5 <- countydemo %>%  filter (pop12_sqmi < 30) %>% 
                filter (med_age > 30) %>% 
                filter (RatioRenttoOwn > 0.6)
                
```

```{r}
#locate the most recent account of OSHPD funding for projects that are in 
#closure 

hospital_cost_closure <- hospital_cost %>% filter(OSHPD.Project.Status == 
                                                    "In Closure")
```

```{r}
#Optimize data for joining (edit so all county naming is same format)

hospital_cost_closure_clean <- hospital_cost_closure %>% 
                       separate(County, c('Number','County_Name_1', 
                       'County_Name_2', 'County_Name_3' )) 
hospital_cost_closure_clean [is.na(hospital_cost_closure_clean)] <- ""
hospital_cost_closure_clean <- hospital_cost_closure_clean %>% unite(County,
                              'County_Name_1','County_Name_2','County_Name_3', 
                              sep = " ", remove = TRUE ) 

county_demo_clean <- countydemo %>% separate(County, c('County_Name_1', 
                      'County_Name_2', 'County_Name_3' )) 
county_demo_clean [is.na(county_demo_clean )] <- ""
county_demo_clean <- county_demo_clean %>% 
                      unite(County,'County_Name_1','County_Name_2',
                      'County_Name_3', sep = " ", remove = TRUE )

mortality_chronic_clean <- mortality_chronic %>% separate(County, c(
                      'County_Name_1','County_Name_2', 'County_Name_3' )) 
mortality_chronic_clean [is.na(mortality_chronic_clean)] <- ""
mortality_chronic_clean <- mortality_chronic_clean %>%
                      unite(County,'County_Name_1','County_Name_2',
                      'County_Name_3', sep = " ", remove = TRUE ) 

```

# Milestone 4 Visualizations 

## Join Datasets for Analysis 
```{r}
Merged_Dataset <- left_join(hospital_cost_closure_clean, county_demo_clean, 
                  by = "County", copy = TRUE)
Merged_Dataset <- left_join(Merged_Dataset, mortality_chronic_clean, 
                  by = "County", copy = TRUE)
```

## Figure 1: Plot of Requested Population Parameters
```{r, echo = TRUE}
library(ggplot2)

ggplot(data = county_demo_clean) + geom_point(mapping = aes(x = med_age, y = RatioRenttoOwn,size= pop12_sqmi),color = "black", fill = "black"
)+theme_bw()+labs(x = "Median Age", y = "Ratio Renters to Owners", title =
"Median Age vs. Renters to Owners vs. Population Size", subtitle = 
"Represented at the County Level", size = "Populattion per SqMi")
```

Interpretation: County Level Data indicates that Median Age has an inverse relationship with proportion of Renters: older populations have a higher 
proportion of home ownership. The largest populations tend to sit within the
mode of the dataset in terms of both Median Age and Ratio of Renters to owners. 
\newpage

## Figure 2: Table 
```{r, echo = TRUE}
Merged_Dataset$County <- trimws(Merged_Dataset$County, which = c("right"))

x <- Merged_Dataset %>% filter(County == c("Mendocino", "Mono", "Colusa", "Del Norte", "Glenn"))

x1 <- x %>% group_by(County) %>%
  summarize(total_mortality = sum(Count))

x2 <- left_join(countydemo5, x1, by = "County") %>%
  select(c("County", "pop2012", "pop12_sqmi", "med_age", "owner_occ", "renter_occ", "RatioRenttoOwn", "total_mortality")) %>%
  group_by(County) %>%
  mutate(total_mortality = sum(total_mortality))

kable(x2, digits = 2, booktabs = F, longtable = F, col.names = c("County", "Population", "Population per Square Mile", "Median Age", "Homeowners", "Renters", "Renter to Homeowner Ratio", "Total Mortality"), caption = "2014-2019 Chronic Mortality and Demographic Data of 5 at Risk Counties")

```
Interpretation: Given the 5 counties that we determined to have 3 similar attributes of low population per square mile, high median age, and a high proportion of renters vs. homeowners, we see that the chronic mortality is quite significant in 4/5 counties. Mono County may have no results due to lack of responses (NA inputs) or due to how small the population in the county is. 

\newpage

## Figure 3: Table of Choice 
```{r, echo = TRUE}
x3 <- x %>% group_by(County) %>%
  summarize(total_mortality = sum(Count))

hospital_cost_closure_clean$County <- trimws(hospital_cost_closure_clean$County, which = c("right"))

x4 <- left_join(x3, hospital_cost_closure_clean, by = "County") %>%
  select(c("County", "Total.Costs.of.OSHPD.Projects", "Number.of.OSHPD.Projects", "total_mortality")) %>%
  group_by(County) %>%
  summarize(total_cost = sum(Total.Costs.of.OSHPD.Projects),
         total_projects = sum(Number.of.OSHPD.Projects))

x5 <- left_join(x4, x3, by = "County")
x6 <- left_join(x2, x5, by = "County")

x7 <- x6 %>% select(c("County", "pop12_sqmi", "RatioRenttoOwn", "total_cost", "total_projects", "total_mortality.y"))

kable(x7, digits = 2, booktabs = F, longtable = F, col.names = c("County", "Population per Square Mile", "Renter to Homeowner Ratio", "Total Cost of OSHPD Projects", "Number of OSHPD Projects", "Total Chronic Mortality"), caption = "OSHPD Funding for Projects that are in Closure in 5 Targeted Counties between 2014 and 2019")


```
Interpretation: We can use this chart to make recommendations that these 5 counties would benefit from more funding in OSHPD projects based on the minimal amount they have gotten in the passed and argue it would lessen chronic mortality. 
