---
title: "Milestone 3 Workflow"
author: "Meredith Klashhman and Ruben Vargas"
date: "11/8/2021"
output: pdf_document
---

Connect to Data Files 
```{r}
getwd()

countydemo <- read.csv("/home/rstudio/PHW251_MK_RV/ca_county_demographics.csv",
                       header=TRUE)
mortality <- read.csv("/home/rstudio/PHW251_MK_RV/mort_by_county.csv", header = 
                        TRUE)
hospital_cost <- read.csv("/home/rstudio/PHW251_MK_RV/oshpd_hospital_cost.csv", 
                          header = TRUE)
```


DATA CLEANING 
```{r}
#Edit Mortality to replace NA with 0 

mortality[is.na(mortality)] = 0 
```


NEW VARIABLE
```{r}
#Create a Variable to Indicate Chronic Condition 
 
#Chronic Health Conditions defined as the following: 
    #Alzheimer's disease
    #Chronic lower respiratory diseases
    #Diabetes mellitus
    #Essential hypertension and hypertensive renal disease
    #Chronic liver disease and cirrhosis
    #Parkinson's disease

mortality <- mortality %>% 
  mutate(Chronic = case_when(Cause_Desc %in% c("Alzheimer's disease",
                    "Chronic lower respiratory diseases", "Diabetes mellitus",
                    "Essential hypertension and hypertensive renal disease",
                    "Chronic liver disease and cirrhosis", 
                    "Parkinson's disease") ~ "Yes",
                    TRUE ~ "No"))
```

SUBSET DATA 
```{r}
mortality <- mortality %>% 
              filter(Chronic == "Yes")%>%
              filter (Strata == "Total Population")
```

NEW VARIABLE
```{r}
#Calculate Rent vs Homeowners 

library(dplyr)

countydemo <- countydemo %>% 
              mutate(RatioRenttoOwn = renter_occ/owner_occ)

```

SUBSET DATA 
```{r}
#locate 5 counties that share three common attributes: low population per 
#square mile `pop12_sqmi1`, high median age 
#`med_age`, a high proportion of renters vs. homeowners 

countydemo5 <- countydemo %>%  filter (pop12_sqmi < 30) %>% 
                filter (med_age > 30) %>% 
                filter (RatioRenttoOwn > 0.6)
```

SUBSET DATA 
```{r}

#locate the most recent account of OSHPD funding for projects that are in 
#closure 

hospital_cost_closure <- hospital_cost %>% filter(OSHPD.Project.Status == 
                                                    "In Closure")
```

DATA CLEANING 
```{r}
library(tidyr)

hospital_cost_closure_clean <- hospital_cost_closure %>% 
                       separate(County, c('Number','County_Name_1', 
                       'County_Name_2', 'County_Name_3' )) 
  
hospital_cost_closure_clean [is.na(hospital_cost_closure_clean)] <- ""

hospital_cost_closure_clean <- hospital_cost_closure_clean %>% unite(County_Name,
                              'County_Name_1','County_Name_2','County_Name_3', 
                              sep = " ", remove = TRUE )
```

DATA JOIN 
join the demographics data to your OSHPD funding dataframe.

DATA JOIN 
Like you've done before, you will join the summarized count of Deaths to expand your OSHPD dataset for analysis.

FINAL DOCUMENT 
4.	Data dictionary based on clean dataset (minimum 4 data elements), including:
o	Variable name
o	Data type
o	Description
5.	One or more tables with descriptive statistics for 4 data elements
6.	PDF that is professionally prepared for presentation
o	Each part of the milestone is clearly on one page (use \newpage to push to a new page)
o	Only the necessary information is outputted (you should suppress, for example, entire data frame outputs)
o	Use of headers and sub headers to create an organized document
