---
title: "Milestone 2"
author: "Meredith Klashman and Ruben Vargas"
date: "October 4, 2021" 
output: html_document
---

#For future use, enter into terminal : git config credential.helper store
git config --global user.email "name@example.edu""
git config --global user.name "meredithklashhman"

```{r, include = FALSE}
library(dplyr)
library(tidyverse)
```

Population of Interest: non-homeowners and aging individuals 
- include total count of mortality from chronic health conditions over the past few years
- summarize a total count of occurrences attributed to known chronic health conditions from 2014-2019
- locate 5 counties that share three common attributes: low population per square mile `pop12_sqmi1`, high median age `med_age`, a high proportion of renters vs. homeowners (you may need to create a new variable somehow for this third criteria)
- locate the most recent account of OSHPD funding for projects that are in closure, and join the demographics data to your OSHPD funding dataframe
- expand your dataframe further by exploring the 2014-2019 Final Deaths by Year data, locating the chronic disease counts by their descriptions, grouping and summarizing at the county level, (you may encounter missing values here, it's understood that replacing NAs with 0 is appropriate)
- join the summarized count of Deaths to expand your OSHPD dataset for analysis
- create a print quality table and data visualization (plot)  to bring to the next DHE business strategy meeting and explain why these particular top 5 counties should be selected for development funding partnerships



Description of dataset
What is the data source? (1-2 sentences on where the data is coming from, dates included, etc.)

- The 3 datasets (oshpd hospital cost, mortality by county, & CA county demographics) that we will be utilizing in this project come from the PHW251 Box Folder. For the OSHPD hospital cost we have data dating back to Septemeber 2013, the mortality data from 2014-2019 and the county demographics from 2012. 


How does the dataset relate to the group problem statement and question?

- We will have to manipulate some of the columns and tidy up the data; however, we have all the data that is needed in order to further investigate our problem statement and questions of interest. 


\newpage

Import

Use appropriate import function and package based on the type of file
```{r, import_data}
hospital_cost <- read.csv("/Users/rubenvargas/Desktop/Public Health W251/PHW251_MK_RV/oshpd_hospital_cost.csv")
mort_county <- read.csv("/Users/rubenvargas/Desktop/Public Health W251/PHW251_MK_RV/mort_by_county.csv")
county_demo <- read.csv("/Users/rubenvargas/Desktop/Public Health W251/PHW251_MK_RV/ca_county_demographics.csv")

hospital_cost
mort_county
county_demo
```
Utilize function arguments to control relevant components (i.e. change column types, column names, missing values, etc.)

```{r, data_manipulation}
# will use for future milestones

# here is where we will manipulate and tidy up the data

```


\newpage

Identify data types for 5+ data elements/columns/variables
Identify 5+ data elements required for your specified scenario. If <5 elements are required to complete the analysis, please choose additional variables of interest in the data set to explore in this milestone.

- we will be using `pop2012`, `pop2012_sqmi`, `med_age`, `households`, `owner_occ`, and `renter_occ` from the count_demo dataframe. We will have to use these columns to create a new variable that will be essential for investigating our problem statement. We will also be using race/ethnicity, county, and age factors from the mort_county and county_demo dataframes to narrow down and focus on 5 counties of interest. 

Utilize functions or resources in RStudio to determine the types of each data element (i.e. character, numeric, factor)

```{r, element}
class(county_demo$pop12_sqmi)
typeof(county_demo$pop12_sqmi)

class(county_demo$pop2012)
typeof(county_demo$pop2012)

class(county_demo$med_age)
typeof(county_demo$med_age)

class(county_demo$renter_occ)
typeof(county_demo$renter_occ)

class(county_demo$owner_occ)
typeof(county_demo$owner_occ)

class(county_demo$households)
typeof(county_demo$households)


demo_summary <- summary(county_demo)
cost_summary <- summary(hospital_cost)
mort_summary <- summary(mort_county)

demo_summary
cost_summary
mort_summary
```
Identify the desired type/format for each variable—will you need to convert any columns to numeric or another type? 

- As we do not want to be dealing with "half households" or "0.25 of a person" we likely will keep those as integers and not convert them to numeric. We might actually prefer that all of our values are in integer format and may change the few that are numeric to integer using the `as.integer()` function


Provide a basic description of the 5+ data elements

- We will likely be looking at the means of the 5+ data elements that we have chosen and use the values to determine the 5 counties that we will be focusing on based on the common attributes that were mentioned in the description of our problem statement on page one. We will be focused on looking at the counties with the higher proportion of non-homeowners based on our calculations and manipulation of the data that we are provided. 
