---
title: "Milestone 2"
author: "Meredith Klashman and Ruben Vargas"
date: "September 30, 2021" 
output: html_document
---

# hi meredith
#hi ruben

#For future use, enter into terminal : git config credential.helper store
git config --global user.email "name@example.edu""
git config --global user.name "meredithklashhman"

```{r, include = FALSE}
library(dplyr)
library(tidyverse)
```

Description of dataset
What is the data source? (1-2 sentences on where the data is coming from, dates included, etc.)
How does the dataset relate to the group problem statement and question?

Import statement
NOTE: Please use datasets available in the PHW251 Box folder (Links to an external site.) (this is important to make sure everyone is using the same datasets)
```{r, import_data}
hospital_cost <- read.csv("/Users/rubenvargas/Desktop/Public Health W251/PHW251_MK_RV/oshpd_hospital_cost.csv")
mort_county <- read.csv("/Users/rubenvargas/Desktop/Public Health W251/PHW251_MK_RV/mort_by_county.csv")
county_demo <- read.csv("/Users/rubenvargas/Desktop/Public Health W251/PHW251_MK_RV/ca_county_demographics.csv")

```
Use appropriate import function and package based on the type of file
Utilize function arguments to control relevant components (i.e. change column types, column names, missing values, etc.)
Document the import process

Identify data types for 5+ data elements/columns/variables
Identify 5+ data elements required for your specified scenario. If <5 elements are required to complete the analysis, please choose additional variables of interest in the data set to explore in this milestone.
Utilize functions or resources in RStudio to determine the types of each data element (i.e. character, numeric, factor)
Identify the desired type/format for each variable—will you need to convert any columns to numeric or another type? 

Provide a basic description of the 5+ data elements
Numeric: mean, median, range
Character:  unique values/categories
Or any other descriptives that will be useful to the analysis
